When are Fuzzy Extractors Possible?

### 2014-10-06 trunk/info-bounds/FuzzyExtPossibility.tex
couple last minute modifications

\section{Impossibility of Fuzzy Extractors for all Families with $\Hfuzz$}
\label{sec:imposs fuzz ext}
In the previous section, we showed a family of distributions that does not admit a secure sketch.  We provide a similar result for fuzzy extractors.

@@

Our counterexample uses a slightly different family of distributions $\mathcal{W}$ than the counterexample for secure sketches.  
We can't work directly on the Hamming distance over a large alphabet~(as we did for secure sketches). \lnote{it would be good to explain why not}  We embed this larger space to the binary Hamming metric.
Let $x_1,..., x_\nu \in \zo^\nu$.  Let $\mathbb{F}$ denote the field of size $2^{\nu}$.  Let $a_2,..., a_{n/\nu}\in\mathbb{F}$ such that $a_i\neq 0$ and let $b_2,..., b_{n/\nu}\in\mathbb{F}$.  
Interpret $x_1,..., x_{\nu}$ as a element $x\in \mathbb{F}$ and let 
\[w =  \begin{pmatrix} \vec{1} \\a_2  \\ \vdots \\ a_{n/\nu} \end{pmatrix} x + \begin{pmatrix} 0  \\ 
b_2\\ \vdots \\ b_{n/\nu} \end{pmatrix} .
\]
The multiplication is in $\mathbb{F}$.
Define a distribution $W$ as the uniform distribution over values of $x$ for a particular value of $a_2,..., a_{n/\nu}$, $b_2,..., b_{n/\nu}$.  
Let $\mathcal{W}$ be the set of all such $W$.

@@

%The difference between these two games is which player announces their action first, in $\Exp^{\mathcal{W}}_1$, the challenger announces a pair of algorithms and the adversary specifies a source.  In $\Exp^{\mathcal{W}}_2$, the adversary specifies a distribution $V$ over sources, and the challenger then specifies their algorithm.  
This means that showing security for a family of distributions $\mathcal{W}$ is equivalent to showing security for all  distributions $V$ when the distribution is known to the algorithms $V$.  In our negative results, the adversary  uses the uniform distribution $V$ over $\mathcal{W}$.  However, it may be possible to improve parameters by using a different $V$. This would just rule out some member of $\mathcal{W}$ not an average member.  This is true for fuzzy extractors as well and holds changes in parameters including imperfect correctness.

@@

The best strategy for an adversary that receives $i$ as is to guess a point that has the most nearby weight in that level.  The adversary chooses \[w^*= \argmax_{w' \in \mathcal{M}}\Pr_{w\in W | 2^{-(i+1)}< \Pr[W=w]\le 2^{-i}\wedge \dis(w, w^*)}[W=w].\] The success of this adversary is at least $2^{-(i+1)}|B_{t,i, max}\cap\supp(W)|$ as there at $|B_{t, i, max}\cap\supp(W)|$ nearby points in that layer each with probability at least $2^{-(i+1)}$.  There are $n$ outcomes for $i$. The overall success of such an adversary is at most $n$ better than an adversary without such input~(by~\cite[Lemma 2.2]{DBLP:journals/siamcomp/DodisORS08}).  That is, 
\begin{align}
\expe_{i | m\le i \le m+n}2^{-(i+1)}|B_{t, i, max}\cap\supp(W)|&\le \expe_{i | m\le i \le n+m}\left( \max_{w^*\in W}\sum_{w\in W| 2^{-(i+1)}< \Pr[W=w]\le 2^{-i} \wedge \dis(w, w^*) \le t}\Pr [W=w]\right) \nonumber\\
&\le n \left(\max_{w^*\in W} \sum_{w\in W | \dis(w, w^*)\le t} \Pr[W=w]\right)\nonumber \\
&= n 2^{-\Hfuzz(W)}\label{eq:link fuzz 3}
\end{align}
\textbf{Security:}
We now argue security.  First note that the total weight of points whose probability is less than $2^{-(n+m)}$ is at most $2^{-m}$~(there are at most $2^n$ points in the distribution).  Let $1_{\text{low}}$ be the indicator random variable for $\Pr[W=w]\le 2^{-(n+m)}$.  Then 
\begin{align*}
\Hav(W | \sketch(W)) = -\log \left(\Pr[1_{\text{low}}=1] * 1 + \Pr[1_{\text{low}} =0]   2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\right)\\
-\log\left( 2^{-m} + (1-2^{-m})2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\right)
\end{align*}
For the remainder of the proof, we seek a bound on 
\[
2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} =0} = \max_{w\in W | 2^{-(n+m)}<\Pr[W=w]}\Pr[W=w | \sketch(W)].
\]
We separate out this quantity into levels:
\begin{align*}
\max_{w\in W | \Pr[W=w]>2^{-(m+n)}}\left(\Pr[W=w | \sketch(W)]\right) &= \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w | \sketch(W), i]\right)\\
&= \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w]\cdot 2^{|\sketch(W)|i|}\right)\\
&\le \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w]\cdot 2^{H_0(\sketch(W) | i)}\right)\\
&\le \expe_{i | m\le i \le m+n} \left(2^{-i}*|B_{t, i, max}\cap\supp(W)|/\delta\right)\\
&\le\frac{ \expe_{i | m\le i \le m+n} \left(2^{-(i+1)}\cdot |B_{t, i, max}\cap \supp(W)|\right)}{2\delta}\\
%&=\delta \sum_{i | m\le i \le m+n} \Pr[W\in L_i]\left(2^{-i}*|B_{t, i, max}|\right)
%\end{align*}
%Now consider This allows us to conclude that 
%\begin{align*}
%\max_{w\in W | \Pr[W=w]>2^{-n}\epsilon}\left(\Pr[W=w | \sketch(W)]\right) 
&= \frac{n 2^{-\Hfuzz(W)}}{2\delta}.
\end{align*}
Where the last line follows by Equation~\eqref{eq:link fuzz 3}.
Combining both cases we have:
\begin{align*}
\Hav(W | \sketch(W)) &= -\log \left(2^{-m}+\frac{(1-2^{-m})(n)2^{-\Hfuzz(W)}}{2\delta}\right)\\
&\ge -\log \min\{2^{-m}, \frac{(1-2^{-m}) n2^{-\Hfuzz(W)}}{2\delta}\})-1\\
&\ge \Hfuzz(W) - \log n + \log \delta - \log (1-2^{-m}) - 2\\
&\ge \Hfuzz(W) - \log n + \log \delta - 3\\
\end{align*}
Where the third line follows from the second because $\Hfuzz(W)\le \Hoo(W) = m$. The last line follows from the fourth because if $m\ge 1$ then $\log (1-2^{-m})\le 1$ and if $m< 1$ the entire bound is vacuous as $\Hfuzz(W)< 1$.
\end{proof}

@@

\begin{lemma}
\label{lem:fuzz can't get key}
Fix some $(\gen, \rep)$ algorithm with $\kappa \ge 2$.  There exists an information-theoretic distinguisher between $(R, P, Z)$ and $(U_\kappa, P, Z)$ with advantage $\epsilon = 1/4-\ngl(n)$.
\end{lemma}
\begin{proof}
As in the proof of \thref{thm:imposs sketch}, we assume that $\rep$ is deterministic.  Denote by $(\Key, P) \leftarrow \gen(V)$.
By Markov's inequality, there exists a set $A_{p}$ such that $\Pr[p\in A_{p}]\ge 1/2$ and $\forall p\in A_{p}$, 
\[
(\Key |P =p, P = p ) \approx_{2\epsilon} (U_\kappa , P =p).
\]
%\{w' | \dis (w', w)\le t \wedge \rec(w', p) \neq w\}\le 2\delta < 1/2.\]

>>>>

\section{Impossibility of Fuzzy Extractors for a Family with $\Hfuzz$}
\label{sec:imposs fuzz ext}
In the previous section, we showed a family of distributions that does not admit a secure sketch.  We provide a similar result for fuzzy extractors.

@@

Our counterexample uses a slightly different family of distributions $\mathcal{W}$ than the counterexample for secure sketches.  
We will work over a binary alphabet~(we used a large alphabet in our counterexample for secure sketches).  A  property of the binary Hamming space is that a large fraction of any set of bounded size is the near ``boundary'' of that set.  This will be crucial in our proof.  We will embed the larger alphabet we used into the binary Hamming metric.
Let $x_1,..., x_\nu \in \zo^\nu$.  Let $\mathbb{F}$ denote the field of size $2^{\nu}$.  Let $a_2,..., a_{n/\nu}\in\mathbb{F}$ such that $a_i\neq 0$ and let $b_2,..., b_{n/\nu}\in\mathbb{F}$.  
Interpret $x_1,..., x_{\nu}$ as a element $x\in \mathbb{F}$ and let 
\[w =  \begin{pmatrix} \vec{1} \\a_2  \\ \vdots \\ a_{n/\nu} \end{pmatrix} x + \begin{pmatrix} 0  \\ 
b_2\\ \vdots \\ b_{n/\nu} \end{pmatrix} .
\]
The multiplication is in $\mathbb{F}$.
Define a distribution $W$ as the uniform distribution over values of $x$ for a particular value of $a_2,..., a_{n/\nu}$, $b_2,..., b_{n/\nu}$.  
Let $\mathcal{W}$ be the set of all such $W$.

@@

%The difference between these two games is which player announces their action first, in $\Exp^{\mathcal{W}}_1$, the challenger announces a pair of algorithms and the adversary specifies a source.  In $\Exp^{\mathcal{W}}_2$, the adversary specifies a distribution $V$ over sources, and the challenger then specifies their algorithm.  
This means that showing security for a family of distributions $\mathcal{W}$ is equivalent to showing security for all  distributions $V$ when the distribution is known to the algorithms $V$.  In our negative results, the adversary  uses the uniform distribution $V$ over $\mathcal{W}$.  However, it may be possible to improve parameters by using a different $V$. This would just rule out some member of $\mathcal{W}$ not an average member.  This is true for fuzzy extractors as well and is resilient to changes in parameters including imperfect correctness.

@@

The best strategy for an adversary that receives $i$ as is to guess a point that has the most nearby weight in that level.  The adversary chooses \[w^*= \argmax_{w' \in \mathcal{M}}\Pr_{w\in W | 2^{-(i+1)}< \Pr[W=w]\le 2^{-i}\wedge \dis(w, w^*)}[W=w].\] The success of this adversary is at least $2^{-(i+1)}|B_{t,W_i, max}|$ as there at $|B_{t,W_i,max}|$ nearby points in that layer each with probability at least $2^{-(i+1)}$.  There are $n$ outcomes for $i$. The overall success of such an adversary is at most $n$ better than an adversary without such input~(by~\cite[Lemma 2.2]{DBLP:journals/siamcomp/DodisORS08}).  That is, 
\begin{align}
\expe_{i | m\le i \le m+n}2^{-(i+1)}|B_{t,W_i,max}|&\le \expe_{i | m\le i \le n+m}\left( \max_{w^*\in W}\sum_{w\in W| 2^{-(i+1)}< \Pr[W=w]\le 2^{-i} \wedge \dis(w, w^*) \le t}\Pr [W=w]\right) \nonumber\\
&\le n \left(\max_{w^*\in W} \sum_{w\in W | \dis(w, w^*)\le t} \Pr[W=w]\right)\nonumber \\
&= n 2^{-\Hfuzz(W)}\label{eq:link fuzz 3}
\end{align}
\textbf{Security:}
We now argue security.  First note that the total weight of points whose probability is less than $2^{-(n+m)}$ is at most $2^{-m}$~(there are at most $2^n$ points in the distribution).  Let $1_{\text{low}}$ be the indicator random variable for $\Pr[W=w]\le 2^{-(n+m)}$.  Then 
\begin{align*}
\Hav(W | \sketch(W)) = -\log \left(\Pr[1_{\text{low}}=1] * 1 + \Pr[1_{\text{low}} =0]   2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\right)\\
-\log\left( 2^{-m} + (1-2^{-m})2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} = 0)}\right)
\end{align*}
For the remainder of the proof, we seek a bound on 
\[
2^{-\Hav(W | \sketch(W) \wedge 1_{\text{low}} =0} = \max_{w\in W | 2^{-(n+m)}<\Pr[W=w]}\Pr[W=w | \sketch(W)].
\]
We separate out this quantity into levels:
\begin{align*}
\max_{w\in W | \Pr[W=w]>2^{-(m+n)}}\left(\Pr[W=w | \sketch(W)]\right) &= \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w | \sketch(W), i]\right)\\
&= \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w]\cdot 2^{|\sketch(W)|i|}\right)\\
&\le \expe_{i | m\le i \le m+n} \left(\max_{w\in W | \Pr[W=w]\in L_i} \Pr[W=w]\cdot 2^{H_0(\sketch(W) | i)}\right)\\
&\le \expe_{i | m\le i \le m+n} \left(2^{-i}*|B_{t,W_i,max}|/\delta\right)\\
&\le\frac{ \expe_{i | m\le i \le m+n} \left(2^{-(i+1)}\cdot |B_{t,W_i,max}|\right)}{2\delta}\\
%&=\delta \sum_{i | m\le i \le m+n} \Pr[W\in L_i]\left(2^{-i}*|B_{t,W_i,max}|\right)
%\end{align*}
%Now consider This allows us to conclude that 
%\begin{align*}
%\max_{w\in W | \Pr[W=w]>2^{-n}\epsilon}\left(\Pr[W=w | \sketch(W)]\right) 
&= \frac{n 2^{-\Hfuzz(W)}}{2\delta}.
\end{align*}
Where the last line follows by Equation~\eqref{eq:link fuzz 3}.
Combining both cases we have:
\begin{align*}
\Hav(W | \sketch(W)) &= -\log \left(2^{-m}+\frac{(1-2^{-m})(n)2^{-\Hfuzz(W)}}{2\delta}\right)\\
&\ge -\log \min\{2^{-m}, \frac{(1-2^{-m}) n2^{-\Hfuzz(W)}}{2\delta}\})-1\\
&\ge \Hfuzz(W) - \log n + \log \delta - \log (1-2^{-m}) - 2\\
&\ge \Hfuzz(W) - \log n + \log \delta - 3\\
\end{align*}
Where the third line follows from the second because $\Hfuzz(W)\le \Hoo(W) = m$. The last line follows from the fourth because if $m\ge 1$ then $\log (1-2^{-m})\le 1$ and if $m< 1$ the entire bound is vacuous as $\Hfuzz(W)< 1$.
\end{proof}

@@

\begin{lemma}
\label{lem:fuzz can't get key}
Fix some $(\gen, \rep)$ algorithm with $\kappa \ge 2$.  There exists an information-theoretic distinguisher between $(R, P, Z)$ and $(U_\kappa, P, Z)$ with advantage $\epsilon = 1/8-\ngl(n)$.
\end{lemma}
\begin{proof}
As in the proof of \thref{thm:imposs sketch}, we assume that $\rep$ is deterministic.  Denote by $(\Key, P) \leftarrow \gen(V)$.
By Markov's inequality, there exists a set $A_{p}$ such that $\Pr[p\in A_{p}]\ge 1/2$ and $\forall p\in A_{p}$, 
\[
(\Key |P =p, P = p ) \approx_{2\epsilon} (U_\kappa , P =p).
\]
%\{w' | \dis (w', w)\le t \wedge \rec(w', p) \neq w\}\le 2\delta < 1/2.\]
